---
title: "SullyGnome_Srape"
author: "Ran K"
date: "11/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSelenium)
library(rvest)
library(tidyverse)
```

# Introduction and setup.

This project would have not been possible without the amazing 2 part lecture given in **useR! 2018** conference by **Hanjo Odentaal**. <br>
Links - [Part1](https://www.youtube.com/watch?v=OxbvFiYxEzI&t=4284s&ab_channel=RConsortium) , [Part2](https://www.youtube.com/watch?v=JcIeWiljQG4&ab_channel=RConsortium). <br> <br>

In order to correctly scrape using Rselenium and docker - we should install **Docker Toolbox** and **Tight VNC** on our computer, then set up our docker container the command 
1. Lunch docker and set up a chrome enironment, the command is - _"docker run --name chrome -v /dev/shm:/dev/shm -d -p 4445:4444 -p 5901:5900 selenium/standalone-chrome-debug:latest"._ <br>
We should then type the command _"docker ps -a"_ to verify the container is initiated. <br>
Last step is launching Tight VNC and in order to view what our code does. 

# Part 1 - Web Scraping




## Setting up RSelenium and prepering useful functions.

If the instructions were followed correctly, we can set up our RSelenium server and directly navigate ourself into the site we would like to scrape information from.

```{r Web_scraping_prep}
remDr <- remoteDriver(remoteServerAddr = "192.168.99.100", port = 4445L, browser = "chrome")
remDr$open()

Display100Games <- function() {
  # This function uses a dropdown box to display 100 articles per page instead of the default 50
  DropDown_box <- 
    remDr$findElement(using = 'xpath',value = '//*[@id="tblControl_length"]/label/select/option[4]')
  DropDown_box$clickElement()
}

Next_Page <- function() { # This function will navigate press the "next button".
  Next_button <- remDr$findElement(using = 'id',value =  'tblControl_next')
  Next_button$clickElement()
}

snooze <- function() { # Slow and steady is definatly better when you are scraping.
  time <- runif(n = 1, min = 4, max = 6)
  cat("Resting",round(time,digits = 2),"seconds \n")
  Sys.sleep(time)
  
}

Month_List <- #We make a list of months in order to easly string together the month and the year
  c("january","february","march","april","may","june",
    "july","august","september","october","november","december")

df <- data.frame() # This will be our final data frame
```

## Scraping our data

The assemble the database ill preform 2 scraping operations.
1. Scraping historical data regarding each specific game (we choose to scrape 200 games per month). <br>
2. Scraping the general data regarding the total views on twich for a specific month.

### Game specific monthly data

Scraping the javascript table for each specific month using RSelenium and Rvest.

```{r Web_scraping_game_specific}

for (year in c(2016:2020)) {
  for (month in Month_List) {
    
    # First we will load the appropriate page
    remDr$navigate(paste0("https://www.sullygnome.com/games/",year,month,"/watched"))
    cat("Now scraping - \n",remDr$getCurrentUrl()[1][[1]],"\n",sep = "")
    
    snooze()
    
    if (year == 2020 && month == "november") {
      print("Stopping operations")
      break
    } # if
    
    Display100Games() # Just to make sure ill use the function every time we load a page
    
    for (i in seq(1,2)) { 
      # As we will grab 200 games each month, we will need to repeat the process 10 time.
      
     webpage <- read_html(remDr$getPageSource()[[1]])
     Sys.sleep(1.5)
     temp_df <- webpage %>% html_node("table") %>% html_table()
     #temp_df <- temp_df[,-c(1,2,12)] # Removing empty columns.
     temp_df["Date"] <- as.Date(paste("01",month,year), format = "%d %B %Y") # adding a column with the date.
     df <- rbind(temp_df,df) # Binding the temporary dataframe to the full one.
     
     Next_Page()
     snooze()
     
    } # for loop games
  } # for loop month
} # for loop year

# This dataset will need some cleaning, we will save the raw data and clean it later

write.csv(df,"RawData/Twitch_game_data_raw.csv",row.names = FALSE)
```

### General monthly data

Scraping global monthly data using Rvest.

```{r Web_scraping_total}

temp_df <-  # Preparing the dataframe to 
  data.frame(Hours_watched = as.numeric(), Avg_viewers = as.numeric(), Peak_viewers = as.numeric(), 
                   Streams = as.numeric(), Avg_channels = as.numeric(), Games_streamed = as.numeric(),
             Date = as.character())


for (year in c(2016:2020)) {
  for (month in Month_List) {
    
    if (year == 2020 && month == "november") {
      print("Finished")
      break
    }
    
    webpage <- read_html(paste0("https://www.sullygnome.com/",year,month))
    cat("Extracting general data from",month,year,"\n")
    snooze()
    temp <- webpage %>% html_nodes(".InfoStatPanelTLCell") %>% html_text() # Scraping the general information
    temp <- append(temp,paste(month,year)) # Adding the date
    
    temp_row <- (match(year,c(2016:2020))-1)* 12 + match(month,Month_List)
    # Creating a temporary variable to indicate the current row in dataframe:
    # row = (Year -1) * 12 + month 
    for (i in 1:length(temp)) { # We will populate the dataframe one cell at a time
      temp_df[temp_row,i] <- temp[i]
    } # for loop i

  } # for loop month
} # for loop year

write.csv(temp_df,"RawData/Twitch_global_data_raw.csv")
```


## Cleaning the data

We need to do a little bit of cleaning in the "twitch_game_data" file.

```{r Cleaning_data}



df <- read.csv("RawData/Twitch_game_data_raw.csv",header = TRUE)
head(df)
df <- df[,-c(1,12)]
names(df) <- c("Rank","Game","Hours_watched","Hours_Streamed","Peak_viewers","Peak_channels","Streamers","Avg_viewers","Avg_channels","Avg_viewer_ratio","Date")

df$Hours_watched <- gsub(" hours","",df$Hours_watched)
df$Hours_Streamed <- gsub(" hours","",df$Hours_Streamed)

for (column in 3:10) {
  df[,column] <- as.numeric(gsub(",","", df[,column]))
}
df <- df[order(df$Date,df$Rank),]
df <- df[,c(2,1,11,3:10)]
df["Month"] <- sub("\\d{4}-","",df$Date)
df["Month"] <- sub("-01","",df$Month)
df["Year"] <- sub("(?<=\\d{4}).+$","",df$Date,perl = TRUE)
df <- df[,-3]
df <- df[,c(1,2,11,12,3:10)]
write.csv(df,"Datasets/Twitch_game_data.csv",row.names = FALSE)


df2 <- read.csv("RawData/Twitch_global_data_raw.csv",header = TRUE)
df2 <- df2[,-1]
df2["Month"] <- sub(" \\d{4}$","",df2$Date)
df2["Month"] <- match(df2$Month,Month_List)
df2["Year"] <- sub("^\\w+ ","",df2$Date)
df2 <- df2[,-7]
for (column in 1:6) {
  df2[,column] <- as.numeric(gsub(",","", df2[,column]))
}
df2 <- df2[,c(8,7,1:6)]

write.csv(df2,"Datasets/Twitch_global_data.csv",row.names = FALSE)


```




